Simple Wind Models

### 2010-08-02 Winds/papers/Theory/First/wind_simple.tex
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5101 ac11df5e-a788-c23b-a8e1-c758c323924d

@@

As noted in the introduction, there are several reports of line
emission associated with star-forming galaxies that exhibit cool gas outflows.  
[SHORTEN THE OBSERVER SUMMARY]
[Phillips first reported?]
Bright \ion{Mg}{2} and \ion{Fe}{2}* line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10b}
and \cite{rubin+10c}.   
\cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emisison, both directly and when the authors modeled and
`removed' the $v \approx 0\kms$ absorption component.  Although the
authors suggested the emission could be related to back-scattered
light in the wind, they also allowed that it could be related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rubin+10b} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies. 
[Did Kate see FeII* emission in her stacks? If not, why not?]
[Are there aspects of those works that are now understood, e.g. the
difference in FeII 2586 and FeII 2600]
At high redshift, \ion{Si}{2}* line emission has been reported from
the Lyman break galaxies which was suggested to be
related to \ion{H}{2} regions within the galaxies \citep{shapley03}.
Lastly, \ion{Na}{1} line emission has also been detected in the
stacked spectra of $z \sim 0$ galaxies by \cite{chen}.  
We contend that all of these incidences may be examples of line
scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, our results indicate that line emission could (and
should) be produced by all of these galaxies if the outflows (and the
galaxy emission) are nearly isotropic.

@@

Although there are several effects that will reduce line emission
relative to the absorption produced by the outflow, our analysis
indicates that detectable line emission should be present for [the
majority] of winds.  One of the principal effects of this emission is
that it modifies the observed absorption profiles (e.g.\
Figure~\ref{fig:noemiss}) which may lead to erroneous conclusions
regarding characteristics of the flow.  {\it Analysis of cool gas outflows
that ignores emission may incorrectly lead to conclusions the gas is
partially covered, has a significantly lower peak optical depth,
and/or that a $v \sim 0 \mkms$ ISM component (perhaps even infall) is
absent.}  Figure~\ref{fig:obs_ew} demonstrates a simple observational
consequence of the line emission: reduced estimates for the absorption
equivalent width \ewabs\ of the flow.  In the case of \ion{Mg}{2},
which has the most affected transitions, the \ewabs\ is reduced by
$30-50\%$ for the wind models presented in this paper.  This could
yield a systematically lower optical depth for the wind and therefore
a lower total mass and kinetic energy.  Although not apparent in
Figure~\ref{fig:obs_ew}, the effects are most pronounced for wind
scenarios where the peak optical depth occurs near $v = \mkms$.
Simple geometric projection limits the majority of scattered photon emission to
have $v \approx \pm 200 \mkms$ and the absorption profiles
are filled-in primarily at these velocities.  In
Figure~\ref{fig:obs_peaktau}, we plot the observed peak optical depth
for \mgiia\ 
versus the velocity where $\tau = \tau_{pk}$ for the various wind models (i.e.\
$\tau_{pk}$ vs.\ $v_\tau$ from Tables~\ref{tab:line_diag},
\ref{tab:plaw_diag}).   In every one of the models, the intrinsic peak
optical depth exceeds 10 and occasionally exceeds 100.  For the
majority of cases with $v_\tau < -300\mkm$, one observes $\tau_{pk} <
2$ and would infer the wind is not even optically thick!
The results are similar for \mgiib\ although one infers a somewhat
higher column density because of its lower $f\lambda$ value.
In the majority of cases, one would faslely conclude that hte wind
partially covered the source and also estimate a peak optical depth
lower than the intrinsic.

@@

@@

\section{Summary}
\label{sec:summary}

@@

>>>>

Focuses on the observational consequences and implications if our results.

@@

As noted in the introduction, there are several reports of line
emission associated with star-forming galaxies that exhibit cool gas outflows.  
[SHORTEN THE OBSERVER SUMMARY]
[Phillips first reported?]
Bright \ion{Mg}{2} and \ion{Fe}{2}* line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite{rubin+10b}
and \cite{rubin+10c}.   
\cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emisison, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  Although the
authors suggested the emission could be related to back-scattered
light in the wind, they also allowed that it could be related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rubin+10b} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies. 
[Did Kate see FeII* emission in her stacks? If not, why not?]
[Are there aspects of those works that are now understood, e.g. the
difference in FeII 2586 and FeII 2600]
At high redshift, \ion{Si}{2}* line emission has been reported from
the Lyman break galaxies which was suggested to be
related to \ion{H}{2} regions within the galaxies \citep{shapley03}.
Lastly, \ion{Na}{1} line emission has also been detected in the
stacked spectra of $z \sim 0$ galaxies by \cite{chen}.  
We contend that all of these incidences may be examples of line
scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, our results indicate that line emission could (and
should) be produced by all of these galaxies if the outflows (and the
galaxy emission) are nearly isotropic.

@@

Although there are several effects that will reduce line emission
relative to the absorption produced by the outflow, our analysis
indicates that detectable line emission should be present for [the
majority] of winds.  One of the principal effects of this emission is
that it modifies the observed absorption profiles (e.g.\
Figure~\ref{fig:noemiss}) which may lead to erroneous conclusions
regarding characteristics of the flow.  {\it Analysis of cool gas outflows
that ignores emission may incorrectly lead to conclusions the gas is
partially covered, has a significantly lower peak optical depth,
and/or that a $v \sim 0 \mkms$ ISM component (perhaps even infall) is
absent.}  Figure~\ref{fig:obs_ew} demonstrates a simple observational
consequence of the line emission: reduced estimates for the absorption
equivalent width \ewabs\ of the flow.  In the case of \ion{Mg}{2},
which has the most affected transitions, the \ewabs\ is reduced by
$30-50\%$ for the wind models presented in this paper.  This could
yield a systematically lower optical depth for the wind and therefore
a lower total mass and kinetic energy.  Although not apparent in
Figure~\ref{fig:obs_ew}, the effects are most pronounced for wind
scenarios where the peak optical depth occurs near $v = \mkms$.
Simple geometric projection limits the majority of scattered photon emission to
have $v \approx \pm 200 \mkms$ and the absorption profiles
are filled-in primarily at these velocities.  In
Figure~\ref{fig:obs_peaktau}, we plot the observed peak optical depth
for \mgiia\ 
versus the velocity where $\tau = \tau_{pk}$ for the various wind models (i.e.\
$\tau_{pk}$ vs.\ $v_\tau$ from Tables~\ref{tab:line_diag},
\ref{tab:plaw_diag}).   In every one of the models, the intrinsic peak
optical depth exceeds 10 and occasionally exceeds 100.  For the
majority of cases with $v_\tau < -300\mkms$, one observes $\tau_{pk} <
2$ and would infer the wind is not even optically thick!
The results are similar for \mgiib\ although one infers a somewhat
higher column density because of its lower $f\lambda$ value.
In the majority of cases, one would faslely conclude that hte wind
partially covered the source and also estimate a peak optical depth
lower than the intrinsic.

@@

The previous few paragraphs sound a cautionary and largely negative
perspective on the implications for absorption-line analysis of
galactic-scale outflows, in the presence of (expected) significant
line-emission.   While this does complicate [the analysis],
observations of the line emission alone offer new and unique
constraints on the charcteristics of the outflow.  And when coupled
with the absorption-line data, the two sets of constraints...
The most direct measure provided by the line emission is an estimate
of the size and morphology of the outflow \citep[e.g.][]{rubin_emiss}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths [is that right?].  The principle
challenge is to achieve sufficient sensitivity to detect the low
surface brightness emission.

As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. [Give estimate]
[Add FeII* to the SB figure]

Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for [this analysis] because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  This implies the highest peak and integrated fluxes
(Figure~\ref{fig:obs_sb}).  With an IFU observation covering all of
the \ion{Fe}{2} transitions, one could sum the together to enhance the
signal.

The kinematics of the line emission also offer insight into physical
characteristics of the wind.

@@

\section{Concluding remarks}
\label{sec:conclude}

1. More complicated winds needed

2. Cosmological context (espeicially at $z>2$)

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{Figures/fig_obs_sb.ps}
\caption{
Surface brightness
}
\label{fig:obs_sb}
\end{figure}
